---
title: "ps_7"
author: "Arnav Srivastava"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# loading relavent libraries for data processing and cleaning

library(tidyverse)
library(skimr)
library(readr)
library(gt)
library(broom)


# reading data set NOMINATE, which is in raw-data folder

data <- read_csv("raw-data/HSall_members.csv")
```

## Question 1: Conduct exploratory data analysis

1A) Summary Statistics
```{r p1, echo = FALSE}

# filtering data for both chambers of 116th congress, looking at only Dems and
# Reps. I use map to put a name to whether a congressman is Dem or Rep (but
# since I used map, I needed to unnest the list in order to see the values). I
# then calculated age of each congressman (subtracted birth year from 2019) and
# their percentile of nominate_dim1 (using function percent_rank and multiplying
# by 100 so that '50 is someone at the 50th percentile'). Finally, I selected
# for desired cols and used skim()

modern <- data %>% 
  filter(chamber == "Senate" | chamber == "House") %>% 
  filter(congress == 116) %>% 
  filter(party_code == 100 | party_code == 200) %>% 
  mutate(party = map(party_code, ~ ifelse(. == 100,
                                          "Democrat",
                                          "Republican")
                     )
  ) %>%
  unnest(party) %>% 
  mutate(age = 2019 - born) %>% 
  mutate(nominate_percentile = percent_rank(nominate_dim1)*100)

modern %>% 
  select(nominate_percentile, age) %>% 
  skim()
  

```

1B) Visualizing a single variable 

```{r p2, echo = FALSE}

# made a plot exploring age distribution of congressmen in 116th congress,
# facetting data by party to see if there is a difference between age
# distribution between the parties (giving insight to whether there are more
# older people in the more conservative Republican party). The graphic shows
# that the age demographics for both parties is more or less similar, but the
# Republican party seems to have more people in the higher-aged bars of the
# histogram

ggplot(modern, aes(x = age, fill = party)) +
  geom_histogram(bins = 15) +
  facet_wrap(~ party) +
  scale_fill_manual(values = c("steelblue", "red")) +
  labs(title = "Age Distribution of 116th Congress",
       subtitle = "Facetted by Democrats and Republicans",
       x = "Age in Years",
       y = "Number of Congressmen", 
       fill = "Party") +
  theme_classic()


# this second plot graphing the nominate_percentile shows that a congressman's
# likelihood of having a fixed conservative ideaology is very strongly
# associated with their party affiliation. The chart also shows that the total
# distribution of Congress is very even, and there is a very slight majority in
# Democratic Party dominance. Finally, this graph shows that there are more
# moderate Democrats at the 50 percentile mark than there are moderate
# Republicans

ggplot(modern, aes(x = nominate_percentile, fill = party)) +
  geom_histogram(bins = 30) +
  scale_fill_manual(values = c("steelblue", "red")) +
  labs(title = "Economic Liberalism-Conservativism of 116th Congress",
       subtitle = "Distribution of economic liberal-conservative in relation to party-affiliation",
       x = "Percentile distribution along Economic Liberalism-Conservatism spectrum",
       y = "Number of Congressmen",
       fill = "Party") 

```

1C) Bivariate Correlations

```{r p3, echo = FALSE}

# used the cor function from summarize to find the linear correlation cofficient
# between age and nominate_percentile cols

modern %>% 
  summarize(correlation = cor(age, nominate_percentile))
```
A negative coefficient indicates that as age increases, the nominate_percentile value for a congressmen tends to go decrease (following a straight line), while a positive coefficient indicates that as age increases, the nominate_percentile value for a congressmen tends to go increase (following a straight line).



1D) Plotting Bivariate relationships

```{r p4, echo = FALSE}

# creates a ggplot showing age vs nominate_percentile. By feeding aes()
# components into each geom function, I was able to plot a black line of best
# fit for the entire dataset while I was able to print a colored line of best
# fit for each party

ggplot(modern, aes(x = age, y = nominate_percentile)) +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE, color = "black") +
  geom_point(aes(color = party)) +
  geom_smooth(aes(color = party), method = "glm", formula = y ~ x, se = FALSE) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Relationship Between Age and DW Nominate Percentile",
       subtitle = "Higher scores represent more conservative positions",
       x = "Age",
       y = "Percentile Rank (DW Nominate)",
       color = "Party") +
  theme_classic()

```


## Question 2: Run a single regression

```{r q2a, echo = FALSE}

# filter modern dataset to democrats only. Then, using
# lm() and tidy(), we construct the linear regression model and use gt() to make
# the table look pretty

dems <- modern %>% 
  filter(party_code == 100) 
  

lm(nominate_percentile ~ age, data = dems) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, conf.low, conf.high) %>% 
  gt() %>% 
  tab_header(title = "Effect of Age on DW Nominate Percentile") %>% 
  tab_spanner(label = "Higher percentile suggests more conservative",
              columns = vars(term, estimate, conf.low, conf.high)
  ) %>% 
  cols_label(term = "Variable",
             estimate = "Estimate",
             conf.low = "Lower bound",
             conf.high = "Higher bound"
             )

```


2B) Interpreting results

The average treatment effect of increasing a congressman's age by 1 year is -0.33: the effect of a democratic member of Congress being one year older is associated with a their percentile rank for DW nominate scores decreasing by 0.33. We should not necessarily interpret the correlation that older democratic senators are less conservative as causal, as there can be confounding factors. For instance, it is possible that more time spent in Congress makes congressmen more liberal; then, age is not the causal factor of making congressmen less conservative, rather it is the number of years a congressman has spent in Congress which affects how liberal they are. Although it is more likely that incumbent democrats will be older, age is then not the causal factor of growing liberalism, but years spent as an incumbent is the causal factor of growing liberalism.
We calculate confidence intervals to determine the strength of our correlation estimate between age and percentile rank. If confidence interval around the estimate was large, then our estimate would be weaker (such that while our estimate is negative, with a large interval upper bound, it is possible that the correlation is actually zero/negative, which changes our interpretation).


2C) Regression and the Rubin Causal Model

We are comparing the potential outcome of conservatism/liberalism for congressmen based on military service (treatment) or no service (control). If the coefficient on military was positive, then the average treatment effect of congress members having served in the military is associated with higher percentile rank for DW nominate scores (more conservative) than control. We can only attribute this difference in potential outcomes to a causal relationship if these results hold in randomized controlled trials, thus accounting for confounding variables and allowing us to truly observe if military service affects liberal/conservative views in congressmen.


## Question 3: Generalize to many regressions

```{r q3a, echo = FALSE}

# create a column which calculates age for every democratic congressman during
# whatever time period of congress they were in by calculating the starting year
# of every congress (where a new congress occurs every 2 years), and then
# subtracting that number from the year they were born. I also reused code for
# nominate_percentile

all <- data %>% 
  filter(chamber == "Senate" | chamber == "House") %>% 
  filter(party_code == 100) %>% 
  mutate(age = (1787 + 2*congress) - born) %>% 
  mutate(nominate_percentile = percent_rank(nominate_dim1)*100) %>% 
  select(congress, chamber, age, nominate_percentile) %>%
  group_by(congress) %>% 
  nest()


# similar to textbook, use map to perform lm() on every nested tibble, extract
# the coeff of correlation for each congress, and the se. Then use the se to
# create confidence intervals around every estimate. Plot using ggplot and
# demonstrate confidence intervals with geom_errorbar

all <- all %>% 
  mutate(mod = map(data, ~ lm(nominate_percentile ~ age, data = .))) %>% 
  mutate(reg_results = map(mod, ~ tidy(.))) %>% 
  mutate(coef = map_dbl(reg_results, ~ filter(., term == "age") %>% pull(estimate))) %>% 
  mutate(se = map_dbl(reg_results, ~ filter(., term == "age") %>% pull(std.error))) %>% 
  mutate(upper = coef + se*1.96) %>% 
  mutate(lower = coef - se*1.96)

ggplot(all, aes(x = congress, y = coef)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(x = congress, ymin = lower, ymax = upper), color = "blue") +
  theme_classic() +
  labs(title = "Average Treatment Effect of Age on Conservatism \nAmong Democrats Over Time",
       subtitle = "Scores use DW Nominate Dimension 1 \n95% Confidence Interval",
       x = "Congress",
       y = "Estimate")


```



Collaborators: None.